---
title: "Data Dashboard for ISSRs in CONUS in 2020"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 2 
    number_sections: false 
    df_print: paged

---

```{r setup, include=FALSE}
```


# 1. Introduction

**Backgrond**

1. Contrail-Cirrus Clouds are estimated to contribute 2% of the Earth‘s total anthropogenic global warming. 
2. Contrail-Cirrus Clouds are formed from the emissions of jet engines under certain atmospheric conditions, like humidity and cold.
2. These atmospheric conditions are known as Ice Super Saturated Regions (ISSR).
3. [ISSRs](https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2012GL051909) are identified as regions with a relative humidity with respect to ice above 100% and a temperature below 233K (to avoid regions of supercooled water clouds). 
4. Studying the volumetric evolution of ISSRs can help to facilitate the design of the air transportation system, thereby helping to prevent climate warming. To be specific, airlines can produce fewer contrail-cirrus clouds by cruising at the time when the volume of ISSRs is low or avoiding crossing ISS regions.


**Datasets**

1. Dataset Description: Rapid Refresh (RAP) Analysis Data with lambert conformal projection of a 20-km resolution. To reduce the cells for analysis, 4 adjacent 20km*20km cells are combined to make a 40-km resolution grid.
2. Dataset Owner: the U.S. National Oceanic and Atmospheric Administration (NOAA) 
3. Dataset Type: open source 
5. Dataset Size: 50GB (hourly updated data in 229 days, all available archive data in 2020) 
6. Dataset Time Range: From 5/17/2020 to 12/31/2020 
7. Dataset Area Range: The contiguous United States (CONUS) 
8. Dataset Access: [HTTPS server](https://www.ncei.noaa.gov/data/rapid-refresh/access/rap-252-20km/analysis/)


**Objective:**

1. Make statistical analysis and visualization for the volumetric evolution of ISSRs in CONUS in 2020.
2. Aims to explore the RAP Rapid Analysis data and find out potential patterns about ISSRs, involving the following aspects:
  - Volumetric evolution of ISSR by hour/day/month
  - Volume of ISSRs at different hours in a day 
  - Ceilings and floors by hour/day/month
  - Volume of ISSRs at different pressure levels 
  - Volume of ISSRs at different latitude/longitude


```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)

# Please modify this directory based on your yes_csv folder
# setwd("D:/George Mason University/5. Spring 2021/DAEN 690/complete_yes_csv")

overall_summary <- read.csv("overall_summary.csv")
overall_summary
layer_summary <- read.csv("layer_summary.csv")
layer_summary
```

# 1. Statistical Analysis for the 3D Grid

## 1.1. Calculate hourly/daily percentage change

```{r echo=FALSE}
# calculate hourly/daily percentage change 
overall_summary <- overall_summary %>%
  mutate(hourly_pct_change = (num_of_ISSR/lag(num_of_ISSR) - 1) * 100,
         daily_pct_change = (num_of_ISSR/lag(num_of_ISSR, 24) - 1) * 100) %>%
  relocate(hourly_pct_change, .before = ceiling) %>%
  relocate(daily_pct_change, .before = ceiling)
overall_summary 
```

## 1.2. Trend of volume_of_ISSR


```{r echo=FALSE}
task2 <- overall_summary %>%
  mutate(month = floor(date/100)%%100)
task2$month<-factor(task2$month)

ggplot(data = task2, mapping = aes(x = datetime_id, y = volume_of_ISSR, fill = month)) +
  geom_bar(stat = "identity") +
  theme( axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  ylab("volume_of_ISSR (%)") + xlab("Hourly Datetime") +
  ggtitle("Trend of hourly updated volume_of_ISSR")

```



  
```{r echo=FALSE, message=FALSE, warning=FALSE}

task3 <- overall_summary %>%
  group_by(date_id, date) %>%
  summarise(avg_volume_of_ISSR = sum(volume_of_ISSR)/24) %>%
  ungroup() %>%
  mutate(pct_change = (avg_volume_of_ISSR/lag(avg_volume_of_ISSR) - 1) * 100) %>%
  mutate(month = floor(date/100)%%100)
task3$date<-factor(task3$date)
task3$month<-factor(task3$month)
task3

ggplot(data = task3, mapping = aes(x = date, y = avg_volume_of_ISSR, fill = month)) +
  geom_bar(stat = "identity") +
  theme( axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  ylab("avg_volume_of_ISSR (%)") + xlab("Dates") +
  ggtitle("Trend of daily average volume_of_ISSR")
      

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
month_days_df = data.frame(month = c(5,6,7,8,9,10,11,12),
                           month_days = c(15,30,31,31,30,31,30,31))
task5 <- overall_summary %>%
  mutate(month = floor(date/100)%%100) %>%
  left_join(month_days_df, by = "month") %>%
  group_by(month) %>%
  summarise(avg_volume_of_ISSR = sum(volume_of_ISSR)/24/min(month_days))

task5$month<-factor(task5$month)
task5

ggplot(data = task5, mapping = aes(x = month, y = avg_volume_of_ISSR, fill= month)) +
  geom_bar(stat = "identity") +
  ylab("avg_volume_of_ISSR (%)") + xlab("Months") +
  ggtitle("Monthly average volume_of_ISSR") +
  geom_text(aes(label = round(avg_volume_of_ISSR, 4)),
            vjust = -1.1, size = 3,
            show.legend = FALSE)
      

```

## 1.3. Midnight and Noon

```{r echo=FALSE}
midnight_noon <- overall_summary %>%
  select(datetime_id, date_id, date, hour, volume_of_ISSR) %>%
  mutate(month = floor(date/100)%%100) %>%
  filter(hour == 0 | hour == 12)
midnight_noon

midnight_noon$hour <- factor(midnight_noon$hour)
ggplot(midnight_noon, aes(x = date, y = volume_of_ISSR, col=hour)) + 
  geom_line() +
  facet_wrap(~ month, nrow = 2,ncol = 4, scales = "free_x") +
  ylab("volume_of_ISSR (%)") + xlab("Dates") +
  ggtitle("Volume_of_ISSR at midnight and noon by month") +
  theme(axis.text.x = element_text(angle = 30))
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
month_days_df = data.frame(month = c(5,6,7,8,9,10,11,12),
                           month_days = c(15,30,31,31,30,31,30,31))

midnight_noon <- overall_summary %>%
  select(datetime_id, date_id, date, hour, volume_of_ISSR) %>%
  mutate(month = floor(date/100)%%100) %>%
  left_join(month_days_df, by = "month") %>%
  filter(hour == 0 | hour == 12) %>%
  group_by(month, hour) %>%
  summarise(avg_volume_of_ISSR = sum(volume_of_ISSR)/month_days)
midnight_noon

midnight_noon$hour <- factor(midnight_noon$hour)
ggplot(midnight_noon, aes(x = month, y = avg_volume_of_ISSR, col=hour)) + 
  geom_line() +
  geom_point() +
  ylab("Monthly average volume_of_ISSR (%)") + xlab("Months") +
  ggtitle("Monthly average volume_of_ISSR at midnight and noon") +
  geom_text(aes(label = round(avg_volume_of_ISSR, 4)),
            vjust = -1.1, size = 3,
            show.legend = FALSE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
month_days_df = data.frame(month = c(5,6,7,8,9,10,11,12),
                           month_days = c(15,30,31,31,30,31,30,31))

midnight_noon <- overall_summary %>%
  select(datetime_id, date_id, date, hour, volume_of_ISSR) %>%
  mutate(month = floor(date/100)%%100) %>%
  left_join(month_days_df, by = "month") %>%
  group_by(month, hour) %>%
  summarise(avg_volume_of_ISSR = sum(volume_of_ISSR)/month_days)
  
midnight_noon

midnight_noon$hour <- factor(midnight_noon$hour)
ggplot(midnight_noon, aes(x = month, y = avg_volume_of_ISSR, col=hour)) + 
  geom_line() +
  ylab("Monthly average volume_of_ISSR (%)") + xlab("Months") +
  ggtitle("Monthly average volume_of_ISSR at each hour")
```



## 1.4. Ceilings and floors by months


```{r echo=FALSE}
ceiling_floor <- overall_summary %>%
  select(datetime_id, date_id, date, hour, ceiling, floor) %>%
  pivot_longer(cols = c(ceiling, floor),
               names_to = 'type',
               values_to = 'pressure_level')
ceiling_floor

ggplot(ceiling_floor, aes(x = datetime_id, y = pressure_level, col=type)) + 
  geom_line() +
  ylab("Pressure Levels (Pa)") + xlab("Hourly Datetimes") +
  ggtitle("Ceilings and Floors at each datetime")+
  scale_y_reverse()

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
ceiling_floor2 <- ceiling_floor %>%
  group_by(date, type) %>%
  summarise(avg_pressure_level = sum(pressure_level)/24) %>%
  ungroup() %>%
  mutate(month = floor(date/100)%%100) 

ceiling_floor2

ggplot(ceiling_floor2, aes(x = date, y = avg_pressure_level, col=type)) + 
  geom_line() +
  facet_wrap(~ month, nrow = 2,ncol = 4, scales = "free_x") +
  ylab("Daily Average Pressure Levels (Pa)") + xlab("Dates") +
  ggtitle("Daily average ceilings and floors by month") +
  scale_y_reverse()+
  theme(axis.text.x = element_text(angle = 30))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
month_days_df = data.frame(month = c(5,6,7,8,9,10,11,12),
                           month_days = c(15,30,31,31,30,31,30,31))

ceiling_floor3 <- ceiling_floor %>%
  mutate(month = floor(date/100)%%100) %>%
  left_join(month_days_df, by = "month") %>%
  group_by(month, type) %>%
  summarise(avg_pressure_level = sum(pressure_level)/24/min(month_days))
ceiling_floor3

ggplot(ceiling_floor3, aes(x = month, y = avg_pressure_level, col=type)) + 
  geom_line() +
  geom_point() +
  ylab("Monthly Average Pressure Levels (Pa)") + xlab("Month") +
  ggtitle("Monthly average ceilings and floors") +
  scale_y_reverse() +
  geom_text(aes(label = round(avg_pressure_level, 0)),
            vjust = 1.5, hjust = "inward", size = 3,
            show.legend = FALSE)
```






# 2. Statistical Analysis for Layers/Horizontal Grids

```{r echo=FALSE}
layer_summary
```

## 2.1. Volume of ISSRs (%) at different pressure lavels by months

```{r echo=FALSE, message=FALSE, warning=FALSE}
month_days_df = data.frame(month = c(5,6,7,8,9,10,11,12),
                           month_days = c(15,30,31,31,30,31,30,31))

task4 <- layer_summary %>%
  mutate(month = floor(date/100)%%100) %>%
  left_join(month_days_df, by = "month") %>%
  group_by(month, pressure_level) %>%
  summarise(avg_volume_of_ISSR = sum(volume_of_ISSR)/min(month_days)/24 ) %>%
  arrange(pressure_level)
task4$pressure_level <- factor(task4$pressure_level)
task4$month <- factor(task4$month)
task4

ggplot(data = task4, mapping = aes(x = pressure_level, y =avg_volume_of_ISSR, fill = month)) +
  geom_bar(stat = "identity", position = "dodge") + 
  labs(title = "Volume of ISSRs in different months by pressure levels",
       x = "Pressure Levels (Pa)",
       y = "Volume of ISSRs (%)") +
  theme(axis.text.x = element_text(angle = 45, face="bold"),
        axis.text.y = element_text(face="bold")) +
  theme(plot.title = element_text(size = 15, face="bold.italic"),
        axis.title.x = element_text(face="bold"),
        axis.title.y = element_text(face="bold"))

ggplot(data = task4, mapping = aes(x = month, y =avg_volume_of_ISSR, fill = pressure_level)) +
  geom_bar(stat = "identity", position = "dodge") + 
  labs(title = "Volume of ISSRs in different months by pressure levels",
       x = "Pressure Levels (Pa)",
       y = "Volume of ISSRs (%)") +
  theme(axis.text.x = element_text(face="bold"),
        axis.text.y = element_text(face="bold")) +
  theme(plot.title = element_text(size = 15, face="bold.italic"),
        axis.title.x = element_text(face="bold"),
        axis.title.y = element_text(face="bold"))
```



# 3. Latitude Analysis

```{r include=FALSE}
library(tidyverse)

# Please modify this directory based on your yes_csv folder
setwd("D:/George Mason University/5. Spring 2021/DAEN 690/complete_yes_csv")


latitude_summary <- read.csv("latitude_summary.csv")
latitude_summary
```

This is the latitude summary table. It consists the data of the average volume of the ISSR in whole one year based on each latitude ID. The latitude ID ranges from 1 to 120. Specifically, the latitude ID represents the area which starts from the south to the North area. The smaller the latitude ID value, the closer the area to the South area. Vice versa, the higher ta value of the latitude ID, the closer it is to the North area. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
latitude_year <- latitude_summary %>%
  group_by(lat_id) %>%
  summarise(avg_volume_ISSR = sum(count_cells)/(37*150)/24/229*100) 
latitude_year
```

The following picture is the visualization of the summary table. The latitude ID is sorted from 1 to 120 which shows the latitude area from the south to the north area. Through the following picture, it can be seen that there is a trend between the average volume of the ISSR and the Latitude.It shows that the closer the area to the north, the higher the average volume of the ISSR according to the whole one year data.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(latitude_year, aes(x = lat_id, y = avg_volume_ISSR)) + 
  geom_bar(stat = "identity") +
  ylab("avg_volume_ISSR (%)") + xlab("Latitude ID (from 1 to 120)") +
  ggtitle("Yearly avg_volume_ISSR at each latitude")
```

The following table shows the summary of the monthly average volume of the ISSR of each latitude. It contains the months, latitude ID, and the average volume of the ISSR. Specifically, it shows the average volume of the ISSR of each latitude ID from month 5, which is May, to month 12, which is December. The year period of these months is 2020.

```{r echo=FALSE, message=FALSE, warning=FALSE}
month_days_df = data.frame(month = c(5,6,7,8,9,10,11,12),
                           month_days = c(15,30,31,31,30,31,30,31))

latitude_month <- latitude_summary %>%
  mutate(month = floor(date/100)%%100) %>%
  left_join(month_days_df, by = "month") %>%
  group_by(month,lat_id) %>%
  summarise(avg_volume_ISSR = sum(count_cells)/(37*150)/24/min(month_days)*100) 
latitude_month
latitude_month$month <- factor(latitude_month$month)
```

The following picture shows the visualization of the average volume of the ISSR in each latitude on every month. In every month, we can see there is an increment trend of the average volume of the ISSR when it gets closer to the North area. It brings a conclusion that the closer to North area, the higher the average volume of the ISSR in every month.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(latitude_month, aes(x = lat_id, y = avg_volume_ISSR)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~ month, nrow = 2,ncol = 4) +
  ylab("avg_volume_ISSR (%)") + xlab("Latitude ID (from 1 to 120)") +
  ggtitle("Monthly avg_volume_ISSR at each latitude")
```



# 4. Longitude Analysis

```{r include=FALSE}
library(tidyverse)

# Please modify this directory based on your yes_csv folder
setwd("D:/George Mason University/5. Spring 2021/DAEN 690/complete_yes_csv")


longitude_summary <- read.csv("longitude_summary.csv")
longitude_summary
```

This is the summary table of the average volume of the ISSR of each longitude for one year. The table contains the longitude ID and the average volume of the ISSR. The longitude ID ranges from 1 to 150 which represents the higher the value of the longitude ID, the closer the area to the East.

```{r echo=FALSE, message=FALSE, warning=FALSE}
longitude_year <- longitude_summary %>%
  group_by(lon_id) %>%
  summarise(avg_volume_ISSR = sum(count_cells)/(37*112)/24/229*100) 
longitude_year
```

The following picture is the visualization of the summary table above. Again, it shows the trend of the average volume of the ISSR from the West to the East area in one year. According to the following picture, it shows there is a decrement of the average ISSR volume in the middle area between the East and the West area.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(longitude_year, aes(x = lon_id, y = avg_volume_ISSR)) + 
  geom_bar(stat = "identity") +
  ylab("avg_volume_ISSR (%)") + xlab("Longitude ID (from 1 to 150)") +
  ggtitle("Yearly avg_volume_ISSR at each longitude")
```

The following table shows the summary of the average ISSR volume over the months based on each longitude ID. The month ranges from month 5, which is May, to month 12, which is December. The year period of these months is 2020.

```{r echo=FALSE, message=FALSE, warning=FALSE}
month_days_df = data.frame(month = c(5,6,7,8,9,10,11,12),
                           month_days = c(15,30,31,31,30,31,30,31))

longitude_month <- longitude_summary %>%
  mutate(month = floor(date/100)%%100) %>%
  left_join(month_days_df, by = "month") %>%
  group_by(month,lon_id) %>%
  summarise(avg_volume_ISSR = sum(count_cells)/(37*120)/24/min(month_days)*100) 
longitude_month
longitude_month$month <- factor(longitude_month$month)
```

The following picture is the visualization of the average ISSR volume in every month as summarized in the previous table. Accoring to the following picture, the trend from the whole months is hard to conclude as the average ISSR volumes is fluctuating in every month. However, there are tendencies that the average volume is getting lower in the middle, which is between West and East area, from October to November 2021.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(longitude_month, aes(x = lon_id, y = avg_volume_ISSR)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~ month, nrow = 2,ncol = 4) +
  ylab("avg_volume_ISSR (%)") + xlab("Longitude ID (from 1 to 150)") +
  ggtitle("Monthly avg_volume_ISSR at each longitude")
```









