---
title: "Statistical Analysis for ISSRs"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 2 
    number_sections: false 
    df_print: paged

---

```{r setup, include=FALSE}
```

# 0. Preparation

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)

# Please modify this directory based on your yes_csv folder
setwd("D:/George Mason University/5. Spring 2021/DAEN 690/complete_yes_csv")

overall_summary <- read.csv("overall_summary.csv")
overall_summary
layer_summary <- read.csv("layer_summary.csv")
layer_summary
```

# 1. Statistical Analysis for the 3D Grid

## 1.1. Calculate hourly/daily percentage change

```{r echo=FALSE}
# calculate hourly/daily percentage change 
overall_summary <- overall_summary %>%
  mutate(hourly_pct_change = (num_of_ISSR/lag(num_of_ISSR) - 1) * 100,
         daily_pct_change = (num_of_ISSR/lag(num_of_ISSR, 24) - 1) * 100) %>%
  relocate(hourly_pct_change, .before = ceiling) %>%
  relocate(daily_pct_change, .before = ceiling)
overall_summary 
```

## 1.2. Trend of volume_of_ISSR


```{r echo=FALSE}
task2 <- overall_summary %>%
  mutate(month = floor(date/100)%%100)
task2$month<-factor(task2$month)

ggplot(data = task2, mapping = aes(x = datetime_id, y = volume_of_ISSR, fill = month)) +
  geom_bar(stat = "identity") +
  theme( axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  ylab("volume_of_ISSR (%)") + xlab("Hourly Datetime") +
  ggtitle("Trend of hourly updated volume_of_ISSR")

```



  
```{r echo=FALSE, message=FALSE, warning=FALSE}

task3 <- overall_summary %>%
  group_by(date_id, date) %>%
  summarise(avg_volume_of_ISSR = sum(volume_of_ISSR)/24) %>%
  ungroup() %>%
  mutate(pct_change = (avg_volume_of_ISSR/lag(avg_volume_of_ISSR) - 1) * 100) %>%
  mutate(month = floor(date/100)%%100)
task3$date<-factor(task3$date)
task3$month<-factor(task3$month)
task3

ggplot(data = task3, mapping = aes(x = date, y = avg_volume_of_ISSR, fill = month)) +
  geom_bar(stat = "identity") +
  theme( axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  ylab("avg_volume_of_ISSR (%)") + xlab("Dates") +
  ggtitle("Trend of daily average volume_of_ISSR")
      

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
month_days_df = data.frame(month = c(5,6,7,8,9,10,11,12),
                           month_days = c(15,30,31,31,30,31,30,31))
task5 <- overall_summary %>%
  mutate(month = floor(date/100)%%100) %>%
  left_join(month_days_df, by = "month") %>%
  group_by(month) %>%
  summarise(avg_volume_of_ISSR = sum(volume_of_ISSR)/24/min(month_days))

task5$month<-factor(task5$month)
task5

ggplot(data = task5, mapping = aes(x = month, y = avg_volume_of_ISSR, fill= month)) +
  geom_bar(stat = "identity") +
  ylab("avg_volume_of_ISSR (%)") + xlab("Months") +
  ggtitle("Monthly average volume_of_ISSR") +
  geom_text(aes(label = round(avg_volume_of_ISSR, 4)),
            vjust = -1.1, size = 3,
            show.legend = FALSE)
      

```

## 1.3. Midnight and Noon

```{r echo=FALSE}
midnight_noon <- overall_summary %>%
  select(datetime_id, date_id, date, hour, volume_of_ISSR) %>%
  mutate(month = floor(date/100)%%100) %>%
  filter(hour == 0 | hour == 12)
midnight_noon

midnight_noon$hour <- factor(midnight_noon$hour)
ggplot(midnight_noon, aes(x = date, y = volume_of_ISSR, col=hour)) + 
  geom_line() +
  facet_wrap(~ month, nrow = 2,ncol = 4, scales = "free_x") +
  ylab("volume_of_ISSR (%)") + xlab("Dates") +
  ggtitle("Volume_of_ISSR at midnight and noon by month") +
  theme(axis.text.x = element_text(angle = 30))
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
month_days_df = data.frame(month = c(5,6,7,8,9,10,11,12),
                           month_days = c(15,30,31,31,30,31,30,31))

midnight_noon <- overall_summary %>%
  select(datetime_id, date_id, date, hour, volume_of_ISSR) %>%
  mutate(month = floor(date/100)%%100) %>%
  left_join(month_days_df, by = "month") %>%
  filter(hour == 0 | hour == 12) %>%
  group_by(month, hour) %>%
  summarise(avg_volume_of_ISSR = sum(volume_of_ISSR)/month_days)
midnight_noon

midnight_noon$hour <- factor(midnight_noon$hour)
ggplot(midnight_noon, aes(x = month, y = avg_volume_of_ISSR, col=hour)) + 
  geom_line() +
  geom_point() +
  ylab("Monthly average volume_of_ISSR (%)") + xlab("Months") +
  ggtitle("Monthly average volume_of_ISSR at midnight and noon") +
  geom_text(aes(label = round(avg_volume_of_ISSR, 4)),
            vjust = -1.1, size = 3,
            show.legend = FALSE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
month_days_df = data.frame(month = c(5,6,7,8,9,10,11,12),
                           month_days = c(15,30,31,31,30,31,30,31))

midnight_noon <- overall_summary %>%
  select(datetime_id, date_id, date, hour, volume_of_ISSR) %>%
  mutate(month = floor(date/100)%%100) %>%
  left_join(month_days_df, by = "month") %>%
  group_by(month, hour) %>%
  summarise(avg_volume_of_ISSR = sum(volume_of_ISSR)/month_days)
  
midnight_noon

midnight_noon$hour <- factor(midnight_noon$hour)
ggplot(midnight_noon, aes(x = month, y = avg_volume_of_ISSR, col=hour)) + 
  geom_line() +
  ylab("Monthly average volume_of_ISSR (%)") + xlab("Months") +
  ggtitle("Monthly average volume_of_ISSR at each hour")
```



## 1.4. Ceilings and floors by months


```{r echo=FALSE}
ceiling_floor <- overall_summary %>%
  select(datetime_id, date_id, date, hour, ceiling, floor) %>%
  pivot_longer(cols = c(ceiling, floor),
               names_to = 'type',
               values_to = 'pressure_level')
ceiling_floor

ggplot(ceiling_floor, aes(x = datetime_id, y = pressure_level, col=type)) + 
  geom_line() +
  ylab("Pressure Levels (Pa)") + xlab("Hourly Datetimes") +
  ggtitle("Ceilings and Floors at each datetime")+
  scale_y_reverse()

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
ceiling_floor2 <- ceiling_floor %>%
  group_by(date, type) %>%
  summarise(avg_pressure_level = sum(pressure_level)/24) %>%
  ungroup() %>%
  mutate(month = floor(date/100)%%100) 

ceiling_floor2

ggplot(ceiling_floor2, aes(x = date, y = avg_pressure_level, col=type)) + 
  geom_line() +
  facet_wrap(~ month, nrow = 2,ncol = 4, scales = "free_x") +
  ylab("Daily Average Pressure Levels (Pa)") + xlab("Dates") +
  ggtitle("Daily average ceilings and floors by month") +
  scale_y_reverse()+
  theme(axis.text.x = element_text(angle = 30))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
month_days_df = data.frame(month = c(5,6,7,8,9,10,11,12),
                           month_days = c(15,30,31,31,30,31,30,31))

ceiling_floor3 <- ceiling_floor %>%
  mutate(month = floor(date/100)%%100) %>%
  left_join(month_days_df, by = "month") %>%
  group_by(month, type) %>%
  summarise(avg_pressure_level = sum(pressure_level)/24/min(month_days))
ceiling_floor3

ggplot(ceiling_floor3, aes(x = month, y = avg_pressure_level, col=type)) + 
  geom_line() +
  geom_point() +
  ylab("Monthly Average Pressure Levels (Pa)") + xlab("Month") +
  ggtitle("Monthly average ceilings and floors") +
  scale_y_reverse() +
  geom_text(aes(label = round(avg_pressure_level, 0)),
            vjust = 1.5, hjust = "inward", size = 3,
            show.legend = FALSE)
```






# 2. Statistical Analysis for Layers/Horizontal Grids

```{r echo=FALSE}
layer_summary
```

## 2.1. Volume of ISSRs (%) at different pressure lavels by months

```{r echo=FALSE, message=FALSE, warning=FALSE}
month_days_df = data.frame(month = c(5,6,7,8,9,10,11,12),
                           month_days = c(15,30,31,31,30,31,30,31))

task4 <- layer_summary %>%
  mutate(month = floor(date/100)%%100) %>%
  left_join(month_days_df, by = "month") %>%
  group_by(month, pressure_level) %>%
  summarise(avg_volume_of_ISSR = sum(volume_of_ISSR)/min(month_days)/24 ) %>%
  arrange(pressure_level)
task4$pressure_level <- factor(task4$pressure_level)
task4$month <- factor(task4$month)
task4

ggplot(data = task4, mapping = aes(x = pressure_level, y =avg_volume_of_ISSR, fill = month)) +
  geom_bar(stat = "identity", position = "dodge") + 
  labs(title = "Volume of ISSRs in different months by pressure levels",
       x = "Pressure Levels (Pa)",
       y = "Volume of ISSRs (%)") +
  theme(axis.text.x = element_text(angle = 45, face="bold"),
        axis.text.y = element_text(face="bold")) +
  theme(plot.title = element_text(size = 15, face="bold.italic"),
        axis.title.x = element_text(face="bold"),
        axis.title.y = element_text(face="bold"))

ggplot(data = task4, mapping = aes(x = month, y =avg_volume_of_ISSR, fill = pressure_level)) +
  geom_bar(stat = "identity", position = "dodge") + 
  labs(title = "Volume of ISSRs in different months by pressure levels",
       x = "Pressure Levels (Pa)",
       y = "Volume of ISSRs (%)") +
  theme(axis.text.x = element_text(face="bold"),
        axis.text.y = element_text(face="bold")) +
  theme(plot.title = element_text(size = 15, face="bold.italic"),
        axis.title.x = element_text(face="bold"),
        axis.title.y = element_text(face="bold"))
```



# 3. Latitude Analysis

```{r include=FALSE}
library(tidyverse)

# Please modify this directory based on your yes_csv folder
setwd("D:/George Mason University/5. Spring 2021/DAEN 690/complete_yes_csv")


latitude_summary <- read.csv("latitude_summary.csv")
latitude_summary
```

This is the latitude summary table. It consists the data of the average volume of the ISSR in whole one year based on each latitude ID. The latitude ID ranges from 1 to 120. Specifically, the latitude ID represents the area which starts from the south to the North area. The smaller the latitude ID value, the closer the area to the South area. Vice versa, the higher ta value of the latitude ID, the closer it is to the North area. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
latitude_year <- latitude_summary %>%
  group_by(lat_id) %>%
  summarise(avg_volume_ISSR = sum(count_cells)/(37*150)/24/229*100) 
latitude_year
```

The following picture is the visualization of the summary table. The latitude ID is sorted from 1 to 120 which shows the latitude area from the south to the north area. Through the following picture, it can be seen that there is a trend between the average volume of the ISSR and the Latitude.It shows that the closer the area to the north, the higher the average volume of the ISSR according to the whole one year data.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(latitude_year, aes(x = lat_id, y = avg_volume_ISSR)) + 
  geom_bar(stat = "identity") +
  ylab("avg_volume_ISSR (%)") + xlab("Latitude ID (from 1 to 120)") +
  ggtitle("Yearly avg_volume_ISSR at each latitude")
```



```{r echo=FALSE, message=FALSE, warning=FALSE}
month_days_df = data.frame(month = c(5,6,7,8,9,10,11,12),
                           month_days = c(15,30,31,31,30,31,30,31))

latitude_month <- latitude_summary %>%
  mutate(month = floor(date/100)%%100) %>%
  left_join(month_days_df, by = "month") %>%
  group_by(month,lat_id) %>%
  summarise(avg_volume_ISSR = sum(count_cells)/(37*150)/24/min(month_days)*100) 
latitude_month
latitude_month$month <- factor(latitude_month$month)

ggplot(latitude_month, aes(x = lat_id, y = avg_volume_ISSR)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~ month, nrow = 2,ncol = 4) +
  ylab("avg_volume_ISSR (%)") + xlab("Latitude ID (from 1 to 120)") +
  ggtitle("Monthly avg_volume_ISSR at each latitude")

```




# 4. Longitude Analysis

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)

# Please modify this directory based on your yes_csv folder
setwd("D:/George Mason University/5. Spring 2021/DAEN 690/complete_yes_csv")


longitude_summary <- read.csv("longitude_summary.csv")
longitude_summary
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
longitude_year <- longitude_summary %>%
  group_by(lon_id) %>%
  summarise(avg_volume_ISSR = sum(count_cells)/(37*112)/24/229*100) 
longitude_year

ggplot(longitude_year, aes(x = lon_id, y = avg_volume_ISSR)) + 
  geom_bar(stat = "identity") +
  ylab("avg_volume_ISSR (%)") + xlab("Longitude ID (from 1 to 150)") +
  ggtitle("Yearly avg_volume_ISSR at each longitude")


```

```{r echo=FALSE, message=FALSE, warning=FALSE}
month_days_df = data.frame(month = c(5,6,7,8,9,10,11,12),
                           month_days = c(15,30,31,31,30,31,30,31))

longitude_month <- longitude_summary %>%
  mutate(month = floor(date/100)%%100) %>%
  left_join(month_days_df, by = "month") %>%
  group_by(month,lon_id) %>%
  summarise(avg_volume_ISSR = sum(count_cells)/(37*120)/24/min(month_days)*100) 
longitude_month
longitude_month$month <- factor(longitude_month$month)

ggplot(longitude_month, aes(x = lon_id, y = avg_volume_ISSR)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~ month, nrow = 2,ncol = 4) +
  ylab("avg_volume_ISSR (%)") + xlab("Longitude ID (from 1 to 150)") +
  ggtitle("Monthly avg_volume_ISSR at each longitude")

```











